---
title: "R on FHIR example"
output:
  html_document: default
  html_notebook: default
---
Install the RonFHIR library
```{r}
# install.packages("RonFHIR")

library(RonFHIR)
library(dplyr)
library(purrr)
```
a tidy data frame for 50 patients' basic demographic details

```{r}



basic_patient_details <- function(patient){
      tibble(
          name = patient$name[[1]]$family[patient$name[[1]]$use == "official"],
          home_phone = patient$telecom[[1]]$value[patient$telecom[[1]]$use == "home"],
          gender = patient$gender,
          birthDate = as.Date(patient$birthDate),
          city = patient$address[[1]]$city
    )

}

client = fhirClient$new("http://test.fhir.org/r3")
patients = client$read(location = "Patient", summaryType = "json")

patients_df <- patients$entry[[1]]$resource



nrows <- nrow(patients_df)

map_dfr(seq_len(nrows),
        function(x, patients){
        basic_patient_details(patients[x, ])
        }
        , patients = patients_df)

```



Connect to the FHIR API
```{r}
client = fhirClient$new("http://test.fhir.org/r3")
```
Ask for a patient

```{r}
patient = client$read(location = "Patient/example", summaryType = "json")
```
Work with the patient's name data
```{r}
patient_name <- patient$name[[1]]

current_patient_name <- patient_name[patient_name$use == "official",]

paste("Surname:", current_patient_name$family)
paste("Given names:", current_patient_name$given)
paste("Patient UR:",patient$identifier[[1]]$value[1])





```


Lets get a bunch of patients
```{r}
patients = client$read(location = "Patient", summaryType = "json")

```
HOw many patients do i have in R?
```{r}
patients$total
```
Look at the link i.e. helps determine of there is more data to slurp
```{r}
patients$link[[1]]

```
The actual patient data seems to be in a list column called "entry"
in entry is a dataframe
```{r}
head(patients$entry[[1]]$fullUrl)
```
The actual patient records may be in the column call resource
```{r}
patients_df <- patients$entry[[1]]$resource
```

explore the patient_df
```{r}
names(patients_df)
head(patients_df$name)
map(patients_df,typeof)
```


1st patient
```{r}

first_patient <- patients_df[1, ]
#exclude photo
head(first_patient[,-14])
```

```{r}
first_patient_df <- tibble(
name = first_patient$name[[1]]$family[first_patient$name[[1]]$use == "official"],
home_phone = first_patient$telecom[[1]]$value[first_patient$telecom[[1]]$use == "home"],
gender = first_patient$gender,
birthDate = as.Date(first_patient$birthDate),
city = first_patient$address[[1]]$city
)
first_patient_df

```
create a function to pull this dat for each patient into one dataframe
```{r}
basic_patient_details <- function(patient){
      tibble(
          name = patient$name[[1]]$family[patient$name[[1]]$use == "official"],
          home_phone = patient$telecom[[1]]$value[patient$telecom[[1]]$use == "home"],
          gender = patient$gender,
          birthDate = as.Date(patient$birthDate),
          city = patient$address[[1]]$city
    )

}
basic_patient_details(first_patient)
```

Apply this function to our dataset 
```{r}
nrows <- nrow(patients_df)


basic_patient_details(patients_df[2, ])

map_dfr(seq_len(nrows),
        function(x, patients){
        basic_patient_details(patients[x, ])
        }
        , patients = patients_df)
```

Bring it all together

```{r}



basic_patient_details <- function(patient){
      tibble(
          name = patient$name[[1]]$family[patient$name[[1]]$use == "official"],
          home_phone = patient$telecom[[1]]$value[patient$telecom[[1]]$use == "home"],
          gender = patient$gender,
          birthDate = as.Date(patient$birthDate),
          city = patient$address[[1]]$city
    )

}

client = fhirClient$new("http://test.fhir.org/r3")
patients = client$read(location = "Patient", summaryType = "json")

patients_df <- patients$entry[[1]]$resource



nrows <- nrow(patients_df)

map_dfr(seq_len(nrows),
        function(x, patients){
        basic_patient_details(patients[x, ])
        }
        , patients = patients_df)

```







